Manifold fitting by ridge estimation: a subspace-constrained approach

### 2021-03-01 rebuttal/response-referee.tex

\noindent ``\textit {This paper is on manifold learning.  It is very confusing to read.''}\\

{\bf Answer:}\\

To the best of our knowledge, the reserved terminology manifold learning refers to an approach to non-linear dimensionality reduction (Isomap and a number of others), which is a method trying to find the new representation in a lower-dimensional space ($R^d$) by approximately keeping the geometry properties of the input which resides in a high-dimensional ambient space ($R^D$). This line of research is mainly about learning the embedding.

In this paper, we aim to learn the actual manifold in the ambient space. Therefore, our work is not on manifold learning. We call this manifold fitting. There are not many research in this direction. We provided the origin of this line of work in the beginning of paper and also cited the previous of pioneering work from Charles Fefferman et al. 2016, 2018, 2019 with a few others. \\

@@

\noindent ``\textit{For example, realistically the data generated by the manifold model are generated with noise, and the distribution and properties of the noise in conjunction with the reach of the manifold are critical for the recovery. However, these critical issues are bypassed by making strong assumptions whereby only the local behavior of ridges matters.''}\\

{\bf Answer:}\\
We do {\bf not} agree. In fact, in our work, the model in Equation (1) on page 2 assumes the data are generated as 
\begin{equation}\label{model}
x_i=\tilde{x}_i+\epsilon_i, i=1,...,n,
\end{equation}
where $\{\tilde{x}_i, i=1,...,n\}$ are the noiseless signal which belong to some manifold, and $\{\epsilon_i, i=1,...,n\}$ are the i.i.d noise from some distribution, such as the multivariate Gaussian distribution.  We use the observations $x_i$ to recover the manifold. Therefore, we did not bypass consideration of the noise in our model.\\

@@

The reach is a global concept which describes the degree of curvature for the manifold. A larger {``\it reach''} indicates the samples resides further from the manifold can also has a unique projection onto $\cal M$. Therefore, the manifold is more flat globally.\\

We do not prefer to model the properties of the noise in conjunction with the reach in the first place. The reason are two folded:
\begin{itemize}
\item Reach is one way to describe the curvature; if we model the noise in conjunction with the reach, we simply reduce the level of the problem,
\item We only assume the noise (Gaussian) is added to the pure data in the ambient space; this is a more general setup in manifold fitting just like the one in Charles Fefferman et al. 2016, 2018, 2019 (see our response to your 4th comment below). 
\end{itemize}

@@

Because of $\mu_s(x,r,h)\in(0,1),\mu_{s,t}(x,r,h)\in (0,1)$, we know 
\[
\mu(r,h) = \sup_x \max\{\max_s \mu_s(x,r,h), \max_{s,t}\mu_{s,t}(x,r,h)\}\in [0,1].
\]

Assumption 3.4 is a bit stronger by requiring $\mu(r,h)$ to be upper-bounded by a number $\ell$ which is less than 1 for all $h$. This assumption can be neglected and discussed depending on the different occasions of the upper-bound of $\mu(r,h)$. \\

\noindent ``\textit{The simulations are too limited as there are many other competing methods to recover the manifolds that are considered, and these manifolds are too simple to be of much interest.'' }\\

{\bf Answer:}\\
We almost compared all methods that we know of in this filed on the one-dimensional manifold to show the good recovery property of our method under the criteria of average margin and Hausdorff distance. The manifold fitting results (theory and numerical results) do not rely on the data set too much. Note that we coded the abstract fitting algorithm from  Charles Fefferman et al. 2016, 2018, 2019 as well.

@@

In our revised paper, we have added another experiment corresponding to a 2-dimensional sphere in the 3D ambient space. We can add more experiments by increase the $D$ and $d$. We choose not to do so since we have decided to  add a real data analysis where we roughly know the truth, to support our theory instead.

Our method can be applied to general input of data with a manifold hypothesis. We used the above two cases since we do know the manifold hypothesis is true in the data and we completely know the pure manifold so that we can evaluate the distance for comparison.

### 2021-03-01 Ridge_Estimation_main/Ridge-final.tex
zz

There has also been some theoretical research into manifold learning using hypothesis testing. For instance, \cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. Mohammed and Narayanan proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function (asdf) is orthogonal to the subspace spanned by the largest $D-d$ eigenvectors of the Hessian of the asdf \citep{mohammed2017manifold}.

@@

Overall, the problem of manifold estimation is treated as a statistical manifold-fitting problem under the observed data $\{x_i, i=1:n\}$. The observed data, it may be assumed, are constructed as 
\begin{equation}\label{model_ass}
x_i=\tilde{x}_i+\epsilon_i, i=1:n,
\end{equation}
where noiseless data $\{\tilde{x}_i, i= 1:n\}$ are assumed to be sampled from some unknown manifold $\mathcal M$ and the noise $\epsilon_i$ obeys the rules of some distributions, such as the Gaussian distribution.  %With these observations, we can solve the manifold-interpolation problem to pull an outlier $x$ back to the manifold.  
Thus, the manifold-estimation problem is aimed at recovering a set of new samples $\{\hat{x}_i, i= 1:n'\}$, which we label $\hat{\cal G}$ from the data $\{x_i\}$ such that the recovered samples in $\hat{\cal G}$ stay in the real manifold $\mathcal M$ as much as possible. Normally, we can obtain $\hat{\cal G}$ by constructing an estimator using the data:
\begin{equation}\label{Estimator}
\hat{\cal G} = \hat{\cal G}(x_1,...,x_n | \Theta),
\end{equation}
where $\Theta$ represents the parameters used in the estimator. It is worth noting that the number of recovered samples in $\hat{\cal G}$ may be different from that of the data because we can always resample to find more points. Here, $\hat{\cal G}$ is named as a ridge estimator (discrete set). Regardless of the methods/algorithms used to obtain $\hat{\cal G}$, this estimator is, in principle, a subset of the ridge (continuous function) defined for the manifold $\cal{M}$; the connections between them will be elaborated in the later sections.

>>>>

There has also been some theoretical research into manifold learning using hypothesis testing \citep{fefferman2016testing,fefferman2018fitting,fefferman2019reconstruction}. For instance, \cite{fefferman2016testing} developed an algorithm that tests the manifold hypothesis and
\cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. Mohammed and Narayanan proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function (asdf) is orthogonal to the subspace spanned by the largest $D-d$ eigenvectors of the Hessian of the asdf \citep{mohammed2017manifold}.

@@

Overall, the problem of manifold estimation is treated as a statistical manifold-fitting problem under the observed data $\{x_i, i=1:n\}$. The observed data, it may be assumed, are constructed as 
\begin{equation}\label{model_ass}
x_i=\tilde{x}_i+\epsilon_i, i=1:n,
\end{equation}
where noiseless data $\{\tilde{x}_i, i= 1:n\}$ are assumed to be sampled from some unknown manifold $\mathcal M$ and the noise vectors $\{\epsilon_i, i= 1:n\}$ are i.i.d and each $\epsilon_i$ obeys the rules of some distributions, such as the multivariate Gaussian distribution supported on ${\mathbb R}^D$, whose density function $f(\epsilon_i)$ yields the form as
\[
f(\epsilon_i)= \frac{1}{{({2\pi\sigma^2)}^{D/2}}}\exp(-\|\epsilon_i\|_2^2/2\sigma^2).
\]%With these observations, we can solve the manifold-interpolation problem to pull an outlier $x$ back to the manifold.  
Thus, the manifold-estimation problem is aimed at recovering a set of new samples $\{\hat{x}_i, i= 1:n'\}$, which we label $\hat{\cal G}$ from the data $\{x_i\}$ such that the recovered samples in $\hat{\cal G}$ stay in the real manifold $\mathcal M$ as much as possible. Normally, we can obtain $\hat{\cal G}$ by constructing an estimator using the data:
\begin{equation}\label{Estimator}
\hat{\cal G} = \hat{\cal G}(x_1,...,x_n | \Theta),
\end{equation}
where $\Theta$ represents the parameters used in the estimator. It is worth noting that the number of recovered samples in $\hat{\cal G}$ may be different from that of the data because we can always resample to find more points. Here, $\hat{\cal G}$ is named as a ridge estimator (discrete set). Regardless of the methods/algorithms used to obtain $\hat{\cal G}$, this estimator is, in principle, a subset of the ridge (continuous function) defined for the manifold $\cal{M}$; the connections between them will be elaborated in the later sections.
